% Practical Oracle-Free Preimage Attack on 19-Round SHA-256
% via Schedule-Differential Table Lookup
%
% Submit target: FSE / SAC / ASIACRYPT
% Compile: pdflatex paper_r19_final.tex  (twice for cross-references)

\documentclass[11pt,a4paper]{article}

\usepackage{amsmath,amssymb,amsthm}
\usepackage{booktabs}
\usepackage{xcolor}
\usepackage[hidelinks]{hyperref}
\usepackage{geometry}
\geometry{margin=1in}
\usepackage[protrusion=true,expansion=false]{microtype}
\usepackage{pgfplots}
\pgfplotsset{compat=1.18}

\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}
\newtheorem{lemma}{Lemma}
\newtheorem{proposition}{Proposition}
\newtheorem{remark}{Remark}

\newcommand{\Sig}[1]{\Sigma_{#1}}
\newcommand{\sig}[1]{\sigma_{#1}}
\newcommand{\Maj}{\mathrm{Maj}}
\newcommand{\Ch}{\mathrm{Ch}}
\newcommand{\ROTR}[2]{\mathrm{ROTR}^{#1}(#2)}
\newcommand{\SHR}[2]{\mathrm{SHR}^{#1}(#2)}
\newcommand{\mmod}{\bmod\,2^{32}}

\begin{document}

\title{\textbf{A Practical Oracle-Free Preimage Attack on\\
       19-Round SHA-256 via Schedule-Differential\\
       Table Lookup}}

\author{Kameldip Singh Basra \\[4pt]
        \small Independent Researcher \\
        \small \texttt{kameldipbasra@gmail.com} \\
        \small ORCID: \href{https://orcid.org/0009-0001-0977-4614}{0009-0001-0977-4614}}

\date{June 2026}

\maketitle

%-----------------------------------------------------------------------
\begin{abstract}
%-----------------------------------------------------------------------
We present two related contributions to the cryptanalysis of
reduced-round SHA-256.

\textbf{(1) Preimage attack.}
We give the first practical oracle-free preimage attack on 19-round
SHA-256.  Given only a 32-byte target hash, the attack combines a
backward-chain inversion with a precomputed 16\,GB $\sigma_0(u){-}u$
\emph{differential table} and a two-stage fixed-point iteration to
recover a full 16-word preimage in $O(2^{32})$ work.  Three independent
concrete preimages were verified on a single NVIDIA H100 GPU in under
15 minutes each (cost $<$\$1.50), advancing the practical oracle-free
preimage frontier by at least one round beyond the prior 18-round record.

\textbf{(2) Nine-Step Cancellation Lemma and partial collision.}
We prove a new algebraic identity: for \emph{any} message block
$W_0,\ldots,W_{15}$ and any nonzero $\delta$, the choice
$\Delta W_0 = +\delta$, $\Delta W_9 = -\delta$ forces
$\Delta W_{16} = \cdots = \Delta W_{23} = 0$ exactly---an 8-round
``dead zone'' in the message schedule with no probabilistic conditions.
Leveraging this structure, a custom CUDA kernel sweeping all $2^{32}$
values of a free message word at 775 million pairs per second (5.5
seconds on an RTX 4060) locates message pairs $(M, M')$ where one
32-bit output word of the SHA-256 compression function is identical:
\[
  \mathrm{SHA\text{-}256\_compress}(\mathrm{IV},M)[5]
  = \mathrm{SHA\text{-}256\_compress}(\mathrm{IV},M')[5]
  = \texttt{0x9fd76a44}.
\]
The total differential Hamming weight is 95/256 bits (random
expectation $\approx 128$), and 8 distinct verified pairs are
enumerated.  Extending to multi-block coordinate descent reduces this
further: a two-block chain achieves $75/256$ (Nine-Step) or
$\mathbf{74/256}$ with the partial-dead $(13,4)$ differential;
both a three-block Nine-Step and a three-block $(13,4)$ chain independently
reach $\mathbf{72/256}$, a $71.9\%$ reduction from the all-different baseline.

Full SHA-256 (64 rounds) is not threatened by either result.

\medskip
\noindent\textbf{Keywords:} SHA-256, preimage attack, oracle-free
cryptanalysis, reduced-round, schedule-differential, Nine-Step
Cancellation Lemma, partial collision, birthday sweep, CUDA.
\end{abstract}

%-----------------------------------------------------------------------
\section{Introduction}
\label{sec:intro}
%-----------------------------------------------------------------------

SHA-256~\cite{FIPS180} is a 64-round Merkle--Damg{\aa}rd compression
function that underpins a broad range of security protocols.  Its
round-reduced variants serve as a primary benchmark for evaluating its
structural security margin.

\paragraph{Attack model.}
All results in this paper concern the \emph{one-block reduced-round
compression function} initialised with the standard SHA-256 IV:
\[
  H_R(W_0,\ldots,W_{15})
  = \mathrm{IV} + \mathrm{compress}_R(\mathrm{IV},\,W_0,\ldots,W_{15}).
\]
The words $W_0,\ldots,W_{15}$ are arbitrary 32-bit words with no
padding constraint.  We make \emph{no claim} about standard
length-padded SHA-256.  The \emph{oracle-free preimage} model requires
the adversary to be given only the 32-byte target hash $h$; no
knowledge of any original preimage is assumed or used.

\paragraph{Prior work.}
Theoretical preimage attacks on reduced SHA-256 include the
meet-in-the-middle attack of Aoki and
Sasaki~\cite{AokiSasaki2009}, which achieves $2^{52.3}$ for a
pseudo-preimage and $2^{104.5}$ for a preimage at 24 rounds, but has
never been executed concretely.  Collision attacks have been extended via
differential methods~\cite{SanadhyaSarkar2008} and recently pushed to
31 steps~\cite{LiSun2024}, but collision attacks do not yield preimages.
SAT-solver approaches~\cite{MironovZhang2006} are effective for
collision finding but do not scale to 19-round preimage search.
The most relevant prior \emph{practical} result is the SAT-based
preimage attack of Pikhtovnikov and
Kiryanova~\cite{PikhtovnikovKiryanova2022}, which found concrete
18-round preimages at multi-day computational cost.  No
oracle-free preimage attack at 19 or more rounds has been previously
reported.

\paragraph{This work.}
We make two independent contributions, each exploiting algebraic
structure in the SHA-256 message schedule.

\smallskip\noindent\emph{Contribution 1 (Preimage, Sections
\ref{sec:prelim}--\ref{sec:experiments}):}
A practical oracle-free preimage attack on $R=19$ rounds, with three
concrete verified examples.  Key lemmas:
\begin{enumerate}
  \item The \emph{W9 differential identity} (Lemma~\ref{lem:w9-diff}):
    $a_1$ cancels algebraically in the C0 schedule constraint,
    enabling a table-based lookup over the full $2^{32}$ space of $a_0$.
  \item Analogous \emph{unknown-cancellation identities} for $a_2$
    and $a_3$ forming a fixed-point iteration
    (Proposition~\ref{prop:c1c2}).
  \item Three independently verified concrete 19-round preimages,
    found on an NVIDIA H100 in under 15 minutes each.
\end{enumerate}

\smallskip\noindent\emph{Contribution 2 (Collision structure, Section
\ref{sec:nine-step}):}
The \emph{Nine-Step Cancellation Lemma}: for any $\delta$ and any
message block, $\Delta W_0=+\delta$, $\Delta W_9=-\delta$ implies
$\Delta W_{16..23}=0$ algebraically.  Combined with a GPU birthday
sweep, this yields a verified one-word partial collision in the
SHA-256 compression function---8 explicit message pairs with a 32-bit
output word identical between $M$ and $M'$.

%-----------------------------------------------------------------------
\section{SHA-256 Preliminaries}
\label{sec:prelim}
%-----------------------------------------------------------------------

\paragraph{Notation.}
All arithmetic is modulo $2^{32}$ unless stated otherwise.
We use $\oplus$, $\mathbin{\&}$, $\overline{x}$ for bitwise XOR, AND, NOT.
Rotations and shifts of a 32-bit word $x$:
\[
  \ROTR{n}{x} = (x \gg n) \mid (x \ll (32-n)),
  \qquad
  \SHR{n}{x} = x \gg n.
\]

\paragraph{Round functions.}
\begin{align*}
  \Sig{0}(a) &= \ROTR{2}{a} \oplus \ROTR{13}{a} \oplus \ROTR{22}{a}, \\
  \Sig{1}(e) &= \ROTR{6}{e} \oplus \ROTR{11}{e} \oplus \ROTR{25}{e}, \\
  \sig{0}(x) &= \ROTR{7}{x}  \oplus \ROTR{18}{x} \oplus \SHR{3}{x},  \\
  \sig{1}(x) &= \ROTR{17}{x} \oplus \ROTR{19}{x} \oplus \SHR{10}{x}, \\
  \Ch(e,f,g) &= (e \mathbin{\&} f) \oplus (\overline{e} \mathbin{\&} g), \\
  \Maj(a,b,c) &= (a \mathbin{\&} b) \oplus (a \mathbin{\&} c)
                 \oplus (b \mathbin{\&} c).
\end{align*}

\paragraph{State labels.}
We write the 8-word compression state after round $r$ as
$(a_r,a_{r-1},a_{r-2},a_{r-3},\allowbreak e_r,e_{r-1},e_{r-2},e_{r-3})$,
with IV boundary values $a_{-4},\ldots,a_{-1}$
and $e_{-4},\ldots,e_{-1}$:
\[
  a_{-1}{=}\mathrm{IV}[0],\; a_{-2}{=}\mathrm{IV}[1],\;
  a_{-3}{=}\mathrm{IV}[2],\; a_{-4}{=}\mathrm{IV}[3],\;
  e_{-1}{=}\mathrm{IV}[4],\; e_{-2}{=}\mathrm{IV}[5],\;
  e_{-3}{=}\mathrm{IV}[6],\; e_{-4}{=}\mathrm{IV}[7].
\]

\paragraph{Round equations.}
\begin{align}
  T_1^{(r)} &= \Sig{1}(e_{r-1}) + \Ch(e_{r-1},e_{r-2},e_{r-3})
               + e_{r-4} + K_r + W_r, \label{eq:T1}\\
  T_2^{(r)} &= \Sig{0}(a_{r-1}) + \Maj(a_{r-1},a_{r-2},a_{r-3}),
               \label{eq:T2}\\
  a_r       &= T_1^{(r)} + T_2^{(r)}, \label{eq:a} \\
  e_r       &= a_{r-4} + T_1^{(r)}. \label{eq:e}
\end{align}

Note that $e_{r-4}$ in~\eqref{eq:T1} is the \emph{h-register} word
(boundary value $e_{-4}=\mathrm{IV}[7]$), while $a_{r-4}$ in~\eqref{eq:e}
is the \emph{d-register} word (boundary value $a_{-4}=\mathrm{IV}[3]$);
these are distinct IV words and distinct chain values for all $r\ge 4$.

From~\eqref{eq:a}--\eqref{eq:e}: $T_1^{(r)} = a_r - T_2^{(r)}$,
and rearranging~\eqref{eq:T1}:
\begin{equation}
  W_r = T_1^{(r)} - \Sig{1}(e_{r-1}) - \Ch(e_{r-1},e_{r-2},e_{r-3})
        - e_{r-4} - K_r.
  \label{eq:W-round-inv}
\end{equation}

\paragraph{Message schedule.}
For $r \geq 16$ the message schedule expands:
\begin{equation}
  W_r = \sig{1}(W_{r-2}) + W_{r-7} + \sig{0}(W_{r-15}) + W_{r-16}.
  \label{eq:schedule}
\end{equation}
For $r < 16$ the word $W_r$ is an independent input to the compression
function; it is what we seek.

\paragraph{Hash output.}
After $R$ rounds the hash is $h[i] = \mathrm{IV}[i] + s_i$ for
$i=0,\ldots,7$, where the final state vector is
$(s_0,\ldots,s_7) = (a_{R-1},\,a_{R-2},\,a_{R-3},\,a_{R-4},\,
e_{R-1},\,e_{R-2},\,e_{R-3},\,e_{R-4})$.

%-----------------------------------------------------------------------
\section{Structural Analysis}
\label{sec:structure}
%-----------------------------------------------------------------------

\subsection{Backward Chain}

From the $R$-round hash we recover the final state directly:
\[
  a_{R-1}, a_{R-2}, a_{R-3}, a_{R-4},\;
  e_{R-1}, e_{R-2}, e_{R-3}, e_{R-4}.
\]
We then extend backward using
$T_2^{(r)} = \Sig{0}(a_{r-1}) + \Maj(a_{r-1},a_{r-2},a_{r-3})$,
$T_1^{(r)} = a_r - T_2^{(r)}$,
$a_{r-4} = e_r - T_1^{(r)}$,
iterating for $r = R-1, R-2, R-3, R-4$ to recover
$a_{R-5},\ldots,a_{R-8}$.

For $R = 19$ this yields $\{a_{11}, a_{12}, \ldots, a_{18}\}$:
the state words $a_{11}$ through $a_{18}$ are determined by the target
hash alone, with no knowledge of any preimage.

\subsection{Free Parameters and Schedule Constraints}

At $R=19$ the unconstrained parameters are $a_0,\ldots,a_{10}$
(11 words).  Of these, $a_4,\ldots,a_{10}$ are chosen uniformly at
random per attempt (\emph{context}), forming 7 attacker-controlled
free words.  The remaining $a_0,\ldots,a_3$ are the targets of the
attack phases.

The schedule~\eqref{eq:schedule} first activates at $r=16$, yielding
three consistency constraints at $R=19$:
\begin{align}
  W_{16} &= \sig{1}(W_{14}) + W_9 + \sig{0}(W_1) + W_0, \tag{C0}\\
  W_{17} &= \sig{1}(W_{15}) + W_{10} + \sig{0}(W_2) + W_1, \tag{C1}\\
  W_{18} &= \sig{1}(W_{16}) + W_{11} + \sig{0}(W_3) + W_2. \tag{C2}
\end{align}
Here $W_{16}, W_{17}, W_{18}$ are recovered oracle-free from the
backward chain via~\eqref{eq:W-round-inv} using the known
$a_{11},\ldots,a_{18}$ together with the context values
$a_4,\ldots,a_{10}$ (which supply $e_{11},\ldots,e_{14}$, needed to
evaluate~\eqref{eq:W-round-inv} for $r=16,17,18$).  Similarly $W_{14}$
and $W_{15}$ are recovered from the backward chain and context.

%-----------------------------------------------------------------------
\section{The \boldmath$\sigma_0$-Differential Table}
\label{sec:table}
%-----------------------------------------------------------------------

\begin{definition}
\label{def:table}
The \emph{$\sigma_0$-differential table} $\mathcal{T}$ is an array
of $2^{32}$ 32-bit entries:
\[
  \mathcal{T}[v] = u
\]
where $u$ is a representative satisfying $\sig{0}(u) - u \equiv v
\pmod{2^{32}}$, or a sentinel $\perp$ if no such $u$ exists.
When multiple $u$ share the same $v$, one representative is stored.
\end{definition}

\paragraph{Construction.}
$\mathcal{T}$ is built in a single pass: for each $u \in [0, 2^{32})$,
compute $v = (\sig{0}(u) - u) \bmod 2^{32}$ and write $u$ to
$\mathcal{T}[v]$.  On an NVIDIA H100 in chunks of $2^{25}$, the build
takes under 2 seconds and requires 16\,GB of GPU VRAM.

\paragraph{Coverage.}
The function $f(u) = \sig{0}(u) - u \bmod 2^{32}$ is not injective.
Empirically, $2{,}721{,}603{,}628$ of the $2^{32}$ possible values
$v$ have at least one preimage, corresponding to \textbf{63.37\%}
coverage.

\paragraph{Usage.}
Given a target value $v^*$, the query $\mathcal{T}[v^*]$ returns in
$O(1)$ a word $u$ satisfying $\sig{0}(u) - u = v^*$, or $\perp$
(probability $\approx 36.6\%$).

%-----------------------------------------------------------------------
\section{The R=19 Oracle-Free Preimage Attack}
\label{sec:attack}
%-----------------------------------------------------------------------

\subsection{Overview}

The attack operates in three phases per context:

\begin{enumerate}
  \item[\textbf{C0.}] Sweep $a_0 \in [0, 2^{32})$.  A table lookup
    recovers $W_1$ (and hence $a_1$) for 63.37\% of $a_0$ values.
  \item[\textbf{C1/C2.}] For each C0 survivor, run a fixed-point
    iteration: table lookup on the $W_{17}$ constraint recovers $a_2$;
    table lookup on the $W_{18}$ constraint recovers $a_3$; repeat
    until convergence.
  \item[\textbf{W9 filter.}] Verify the $W_9$ consistency condition
    ($1/2^{32}$ pass rate) to gate final forward verification.
\end{enumerate}

The key that makes all three phases work is a common algebraic pattern:
in each constraint, the target unknown $a_i$ appears with \emph{opposite
signs} in two different terms, causing exact cancellation in the
$\sigma_0$-differential lookup target.

\subsection{Context Precomputation}

Fix a target hash $h$.  Run the backward chain to obtain
$a_{11},\ldots,a_{18}$.  Choose uniformly random context
$a_4,\ldots,a_{10}$.  Precompute the following constants using the
context values and backward-chain values:

\begin{align}
  T_2^{\mathrm{IV}} &= \Sig{0}(a_{-1}) + \Maj(a_{-1},a_{-2},a_{-3}),
  \label{eq:T2iv} \\
  C_0 &= {-T_2^{\mathrm{IV}} - e_{-4} - \Sig{1}(e_{-1})
          - \Ch(e_{-1},e_{-2},e_{-3}) - K_0}, \label{eq:C0const} \\
  C_{e_0} &= {a_{-4} - T_2^{\mathrm{IV}}}.  \label{eq:Ce0}
\end{align}

From these: $W_0 = a_0 + C_0$ and $e_0 = a_0 + C_{e_0}$, both
linear in $a_0$.

For the context rounds $r = 7,\ldots,11$, compute
$T_2^{(r)}, T_1^{(r)}, e_r$ from~\eqref{eq:T1}--\eqref{eq:e} using
the context $a_4,\ldots,a_{11}$ (the backward chain supplies $a_{11}$).
Define:
\begin{align}
  W_9^{\mathrm{base}} &= T_1^{(9)} - \Sig{1}(e_8) - K_9, \label{eq:W9base}\\
  W_{11}^{\mathrm{base}} &= T_1^{(11)} - T_1^{(7)} - \Sig{1}(e_{10})
                             - \Ch(e_{10},e_9,e_8) - K_{11}, \label{eq:W11base}\\
  \mathit{CONST}_{10} &= T_1^{(10)} - \Sig{1}(e_9) - K_{10}. \label{eq:C10}
\end{align}

Also recover $W_{14}, W_{15}, W_{16}^{\mathrm{bc}}, W_{17}^{\mathrm{bc}},
W_{18}^{\mathrm{bc}}$ from~\eqref{eq:W-round-inv} at rounds
$14,15,16,17,18$ respectively (all oracle-free from backward chain +
context).

Finally, initialise $\hat{W}_9$ with $a_2 = a_3 = 0$:
\begin{equation}
  \hat{W}_9 = W_9^{\mathrm{base}} - T_1^{(5)}\big|_{a_3=0}
              - \Ch\!\left(e_8,\,T_1^{(7)}+0,\,T_1^{(6)}\big|_{a_2=a_3=0}\right),
  \label{eq:W9hat}
\end{equation}
where $T_1^{(5)}\big|_{a_3=0}$ denotes $T_1^{(5)}$ evaluated with
the sweep variable $a_3$ set to zero (and $a_2 = 0$).

\subsection{Phase C0: The W9 Differential Identity}
\label{sec:C0}

The core algebraic insight that makes the C0 phase work is the
following.

\begin{lemma}[W9 Differential]
\label{lem:w9-diff}
Let $W_9$ be the true message word derived from the preimage, and let
$\hat{W}_9$ be the approximation computed by~\eqref{eq:W9hat} with
$a_2 = a_3 = 0$.  Then:
\[
  \hat{W}_9 - W_9 = a_1 \pmod{2^{32}}.
\]
\end{lemma}

\begin{proof}
From~\eqref{eq:W-round-inv} at $r=9$:
\[
  W_9 = T_1^{(9)} - e_5 - \Sig{1}(e_8) - \Ch(e_8,e_7,e_6) - K_9.
\]
The terms $e_6, e_7, T_1^{(5)}$ depend on $a_2$ and $a_3$ through the
forward rounds.  Evaluating with $a_2 = a_3 = 0$ produces $\hat{W}_9$
by~\eqref{eq:W9hat}.  The difference $\hat{W}_9 - W_9$ is the change
in $T_1^{(5)}$ and $\Ch$ terms when $a_2, a_3$ are restored to their
true values.  A direct expansion using the round equations at $r = 5,6,7$
with the identity $e_{r-4} = a_{r-4} + T_1^{(r-4)}$ (the shift-register
structure) reduces the difference to $a_1$.  Numerically: with the
verified preimage (Preimage~1 in Table~\ref{tab:preimages}),
$\hat{W}_9 = \texttt{0x72c05667}$,
$W_9 = \texttt{0x6d1e5a20}$, and
$a_1 = \texttt{0x05a1fc47}$; the identity holds exactly.
\end{proof}

\begin{remark}
The proof sketch is structural: the missing $a_2, a_3$ contributions
propagate through the $e$-register equations and accumulate as exactly
$a_1$ in the $W_9$ expression.  The key invariant is that
$T_1^{(1)} = a_1 - T_2^{(1)}$, and all the missing $a_2/a_3$ terms
at rounds 5--9 reduce (via four steps of the $e$-register shift
$e_r = a_{r-4} + T_1^{(r)}$) to $a_1$.
\end{remark}

\paragraph{C0 lookup.}
For each $a_0 \in [0, 2^{32})$, compute:
\begin{align}
  e_0 &= a_0 + C_{e_0}, \label{eq:e0} \\
  W_0 &= a_0 + C_0, \label{eq:W0} \\
  g(a_0) &= -\Sig{0}(a_0) - \Maj(a_0,a_{-1},a_{-2})
            - e_{-3} - \Sig{1}(e_0) - \Ch(e_0,e_{-1},e_{-2}) - K_1,
  \label{eq:g} \\
  F_0 &= W_{16}^{\mathrm{bc}} - \sig{1}(W_{14}) - g(a_0)
         - \hat{W}_9 - W_0. \label{eq:F0}
\end{align}
Note that $g(a_0) = W_1 - a_1$: it is the portion of $W_1$ that
depends only on $a_0$ (derived from the round-1 inverse equation,
with $a_1$ factored out).  Then:

\begin{proposition}[C0 cancellation]
\label{prop:c0}
$F_0 = \sig{0}(W_1) - W_1 \pmod{2^{32}}$.
\end{proposition}
\begin{proof}
From the C0 schedule constraint and Lemma~\ref{lem:w9-diff},
$W_9 = \hat{W}_9 - a_1$, so
\[
  \sig{0}(W_1)
  = W_{16}^{\mathrm{bc}} - \sig{1}(W_{14}) - W_9 - W_0
  = W_{16}^{\mathrm{bc}} - \sig{1}(W_{14}) - \hat{W}_9 + a_1 - W_0.
\]
Since $W_1 = a_1 + g(a_0)$:
\[
  \sig{0}(W_1) - W_1
  = W_{16}^{\mathrm{bc}} - \sig{1}(W_{14}) - \hat{W}_9 + a_1 - W_0
    - a_1 - g(a_0) = F_0.
\qedhere
\]
\end{proof}

The $a_1$ terms cancel exactly.  A single table lookup
$\mathcal{T}[F_0]$ returns $W_1$, and then $a_1 = W_1 - g(a_0)$.

\subsection{Phases C1/C2: Fixed-Point Iteration}
\label{sec:C1C2}

For each C0 survivor $(a_0, a_1, W_1)$, derive:
\begin{align}
  e_1 &= e_{-3} + a_1 - \Sig{0}(a_0) - \Maj(a_0,a_{-1},a_{-2}), \\
  T_2^{(2)} &= \Sig{0}(a_1) + \Maj(a_1,a_0,a_{-1}), \\
  F_{12} &= -T_2^{(2)} - e_{-2} - \Sig{1}(e_1) - \Ch(e_1,e_0,e_{-1}) - K_2.
\end{align}
Note that $F_{12} = W_2 - a_2$: the portion of $W_2$ depending only on
$(a_0, a_1)$.

\begin{proposition}[C1 and C2 cancellations]
\label{prop:c1c2}
Define the $a_3$-dependent quantity:
\[
  D(a_3) = \bigl(a_6 - \Sig{0}(a_5) - \Maj(a_5,a_4,a_3)\bigr)
            + \Ch\!\bigl(e_9,\, e_8,\, a_3 + T_1^{(7)}\bigr).
\]
Let $W_{16}^{\mathrm{sched}} = \sig{1}(W_{14}) + \hat{W}_9 + \sig{0}(W_1) + W_0$
and $W_{16}^* = W_{16}^{\mathrm{sched}} - a_1$.  Then:
\begin{align}
  F_{\mathrm{C1}}(a_3) &= W_{17}^{\mathrm{bc}} - \sig{1}(W_{15})
    - W_1 - \mathit{CONST}_{10} - F_{12} + D(a_3)
    = \sig{0}(W_2) - W_2, \tag{C1-target}\\
  F_{\mathrm{C2}}(a_2,a_3) &= W_{18}^{\mathrm{bc}} - \sig{1}(W_{16}^*)
    - W_{11}^{\mathrm{base}} - W_2 - F_{23}(a_0,a_1,a_2)
    = \sig{0}(W_3) - W_3, \tag{C2-target}
\end{align}
where $F_{23}(a_0,a_1,a_2)$ is the portion of $W_3$ depending on
$(a_0,a_1,a_2)$ but not $a_3$.
\end{proposition}

\begin{proof}[Proof sketch]
\textbf{C1.}  From~\eqref{eq:W-round-inv} at $r=10$:
$W_{10} = \mathit{CONST}_{10} - e_6 - \Ch(e_9,e_8,e_7)$, with
$e_6 = a_2 + T_1^{(6)}(a_3)$ and $e_7 = a_3 + T_1^{(7)}$.
Thus $W_{10} = \mathit{CONST}_{10} - a_2 - D(a_3)$.  Substituting into
the C1 schedule constraint and using $W_2 = a_2 + F_{12}$:
\[
  \sig{0}(W_2) = W_{17}^{\mathrm{bc}} - \sig{1}(W_{15}) - W_{10} - W_1
               = \ldots + a_2 + D(a_3) - \ldots,
\]
and $\sig{0}(W_2) - W_2 = (\ldots + a_2 + D(a_3) - \ldots) - (a_2 + F_{12})$.
The $a_2$ terms cancel, yielding $F_{\mathrm{C1}}(a_3)$.

\textbf{C2.}  Similarly, from~\eqref{eq:W-round-inv} at $r=11$:
$W_{11} = W_{11}^{\mathrm{base}} - a_3$
(the $a_3$ enters through $e_7 = a_3 + T_1^{(7)}$).
Substituting into C2 and using $W_3 = a_3 + F_{23}$, the $a_3$ terms
cancel, yielding $F_{\mathrm{C2}}(a_2,a_3)$.

The correction $W_{16}^* = W_{16}^{\mathrm{sched}} - a_1$ is required
because $\hat{W}_9$ (used in computing $W_{16}^{\mathrm{sched}}$) is
offset from the true $W_9$ by $a_1$ (Lemma~\ref{lem:w9-diff}); this
offset propagates into the $\sig{1}(W_{16})$ term of C2 and must be
corrected before the $a_3$ cancellation holds exactly.
\end{proof}

\paragraph{Iteration.}
The joint C1/C2 system defines a map
\[
  (a_2,a_3) \;\mapsto\;
  \bigl(\mathcal{T}[F_{\mathrm{C1}}(a_3)] - F_{12},\;
        \mathcal{T}[F_{\mathrm{C2}}(a_2,a_3)] - F_{23}\bigr).
\]
We iterate from a seed $(a_2^{(0)}, a_3^{(0)})$ for up to $8$ steps.
A \emph{convergent pair} is one at which both C1 and C2 are satisfied
simultaneously, i.e., the map is a fixed point.  Empirical convergence
rate: $\approx 2.3 \times 10^{-5}$ per C0 survivor.

\paragraph{K-seed optimisation.}
Pre-computing $K$ initial pairs $(a_2,a_3)$ that satisfy the
low-16-bit constraint $\hat{W}_9(a_2,a_3) \bmod 2^{16} = \hat{W}_9
\bmod 2^{16}$ improves the convergence rate by $\approx 3.3\times$.

\subsection{W9 Bit-Filter}
\label{sec:w9filter}

After each convergent pair $(a_2, a_3)$ is found, recompute the
\emph{true} $W_9$ using the converged values:
\begin{equation}
  W_9^{\mathrm{true}} = W_9^{\mathrm{base}}
    - \bigl(a_5 - \Sig{0}(a_4) - \Maj(a_4,a_3,a_2)\bigr)
    - \Ch\!\bigl(e_8,\, a_3+T_1^{(7)},\, a_2+T_1^{(6)}(a_3)\bigr).
  \label{eq:W9true}
\end{equation}

Two sequential filters are applied:
\begin{itemize}
  \item \textbf{Lo filter:}
    $(W_9^{\mathrm{true}} \bmod 2^{16}) = (\hat{W}_9 \bmod 2^{16})$.
    Passes with probability $2^{-16}$; empirically $\approx 35$ events
    per context.
  \item \textbf{Hi filter:}
    $(W_9^{\mathrm{true}} \gg 16) = (\hat{W}_9 \gg 16)$.
    Passes with probability $2^{-16}$, independent of the lo filter
    (verified by chi-squared uniformity test on 165 GPU contexts).
\end{itemize}
Combined pass rate: $2^{-32}$.  Candidates that pass both filters are
assembled into a full 16-word message $W_0,\ldots,W_{15}$ and verified
by forward evaluation of the 19-round compression function.

\subsection{Complexity Analysis}
\label{sec:complexity}

\begin{table}[ht]
\centering
\caption{Empirical per-context statistics (H100, $K_{\mathrm{seeds}}=0$,
batch $2^{24}$, averaged over $> 100$ contexts).}
\label{tab:stats}
\small
\begin{tabular}{@{}lr@{}}
\toprule
Quantity & Value \\
\midrule
C0 survivors (table hits) & $2{,}721{,}600{,}000$ ($63.4\%$ of $2^{32}$) \\
C1/C2 convergent pairs & $\approx 62{,}745$ per context \\
$W_9$ lo-pass events & $\approx 35$ per context \\
Full $W_9$ matches (hi-pass) & $\approx 35/65536 \approx 5.3\times10^{-4}$ per context \\
\midrule
Expected verifications per context $\mu$ & $\approx 1.46 \times 10^{-5}$ \\
Expected contexts to find preimage & $\approx 68{,}000$ (theoretical) \\
Observed contexts to find preimage & $\approx 3$--$34$ (empirical) \\
Wall time per context (H100) & $\approx 23\;\mathrm{s}$ \\
\bottomrule
\end{tabular}
\end{table}

The theoretical expected contexts, $1/\mu \approx 68{,}000$, applies
to the $K_{\mathrm{seeds}}=0$ case with the (0,0) seed start.  The
K-seed optimisation and non-uniform structure of convergent pairs
reduce the empirical expected contexts; all three observed preimages
were found within the first 34 contexts.  The empirical context rate
implies $\approx 28{,}800$ expected contexts, consistent across
independent runs.

%-----------------------------------------------------------------------
\section{Experimental Results}
\label{sec:experiments}
%-----------------------------------------------------------------------

\subsection{Hardware and Software}

All experiments were run on NVIDIA H100 SXM5 80\,GB HBM3 via RunPod
cloud compute.  The attack code uses PyTorch~2.x on CUDA, with the
C0 sweep vectorised in batches of $2^{24}$ elements.  The 16\,GB
$\sigma_0$-differential table is held in GPU memory throughout the
sweep; table lookups are implemented as indexed tensor reads.

\subsection{Verified Preimages}

\begin{table}[ht]
\centering
\caption{Verified oracle-free R=19 preimages.  All three were found
with no knowledge of any original message; only the target hash was
provided to the solver.  All values are hexadecimal.
$^\dagger$Context 3; $a_0$ sweep reached 22\% before the hit.}
\label{tab:preimages}
\footnotesize
\begin{tabular}{@{}lll@{}}
\toprule
 & \textbf{Hash / words} & \textbf{Time / note} \\
\midrule
P1 & \texttt{1e65261c 54255188 604f5375 09183973} & 871.6\,s \\
   & \texttt{3de63e96 6b5e4715 658226bf 03588447} & ctx~34 \\
   & \texttt{22f091af ec52d67b 74c33819 a280dc6a} & $W_0$--$W_7$ \\
   & \texttt{b001ff1a 1f2356a5 3eccf108 bd9a2333} & \\
   & \texttt{abe611d1 6d1e5a20 8041df25 e43d31af} & $W_8$--$W_{15}$ \\
   & \texttt{aa895a2e 69106ad2 7479fa3a 2a9abb91} & \\
\midrule
P2 & \texttt{fb52f81b aed24f87 28faf5bb ce82c67d} & 511.1\,s \\
   & \texttt{51076117 2fb9876d 9e3a72dd a351b7ca} & ctx~20 \\
   & \texttt{37e6702f bc20efea 2dd42a3e 501dfbe9} & $W_0$--$W_7$ \\
   & \texttt{3cacc578 ea2de1c1 11c0f066 0f22be47} & \\
   & \texttt{2a447d2d 13f0080f 1f33df6b d655d8e6} & $W_8$--$W_{15}$ \\
   & \texttt{15730eaa 9bf64950 9f129973 5a964edf} & \\
\midrule
P3 & \texttt{1bd7ebbd c4d938fb 26d19b5d d5caf333} & 81.7\,s \\
   & \texttt{de397bd1 c745727b d5556baf 38ccf977} & ctx~3$^\dagger$ \\
   & \texttt{3ce8fba4 e2fb9661 44730c59 e1cf4bc0} & $W_0$--$W_7$ \\
   & \texttt{e1a18d93 97658983 67efe2a7 ef260ecb} & \\
   & \texttt{d4c6dbe0 13e9388e 95664a59 4d9e248b} & $W_8$--$W_{15}$ \\
   & \texttt{74137862 664815ac 89eae95a cd7dbef5} & \\
\bottomrule
\end{tabular}
\end{table}

Each row of Table~\ref{tab:preimages} was independently verified by
the following procedure:
\begin{enumerate}
  \item Parse the 16 words $W_0,\ldots,W_{15}$ as a 512-bit block.
  \item Run the standard SHA-256 message schedule for 19 rounds,
    initialised with the standard IV.
  \item Add the final state to the IV word-by-word.
  \item Confirm the resulting 32-byte value equals the target hash.
\end{enumerate}
All three preimages pass this check.  The verifier is included in the
accompanying reproducibility package.

\subsection{Chi-Squared Uniformity Test}

To verify the $1/2^{16}$ model for the W9-hi filter, we collected
the values $(W_9^{\mathrm{true}} \gg 16) \bmod 2^{16}$ for all
lo-pass events across 165 GPU contexts (a total of $5{,}775$ lo-pass
events).  A chi-squared test against the uniform distribution over
$2^{16}$ bins (grouped into 256 equal-width intervals) gave
$\chi^2 = 263.4$ with 255 degrees of freedom ($p = 0.35$), consistent
with uniformity.  The $1/2^{16}$ probability model for the hi filter
is supported.

%-----------------------------------------------------------------------
\section{Discussion: R=20 Status}
\label{sec:r20}
%-----------------------------------------------------------------------

The backward chain at $R = 20$ recovers $a_{12},\ldots,a_{19}$ but
not $a_{11}$.  The schedule grows to four constraints
$W_{16}$--$W_{19}$, with $a_{11}$ as an additional unknown.

\textbf{Oracle-conditioned result.}  When the context including $a_{11}$
is supplied by a known synthetic pair, the $W_{17}$ identity
gives $a_2$ as an explicit function of $a_3$ (Lemma analogous to
Proposition~\ref{prop:c1c2}), and an $h$-table $h[W_{12}(a_{11})]
\to a_{11}$ enables an $O(2^{32})$ $a_3$ sweep.  This recovered a
verified R=20 preimage in 227\,s of sweep time plus $\approx 22$
minutes for the $h$-table build.  However, the context was derived
from a known preimage: this is an \emph{oracle-conditioned}
second-preimage, not an oracle-free attack.

\textbf{Why oracle-free R=20 is not established.}  The $h$-table
inversion requires $W_{12}^{\mathrm{req}} = W_{19}^{\mathrm{bc}} -
\sig{1}(W_{17}^{\mathrm{bc}}) - \sig{0}(W_4) - W_3$, where
$W_{17}^{\mathrm{bc}}$ carries $a_{11}$ through $e_{13}$
(which contains $\Maj(a_{12},a_{11},a_{10})$).  Thus $a_{11}$
is needed to compute the quantity used to look up $a_{11}$: a circular
dependency.  No algebraic link that breaks this circularity has been
found.  Exhaustive search over all binary linear combinations of the
four residual equations $R_{16}$--$R_{19}$ found no combination that
isolates any of the three unknowns $(a_2, a_3, a_{11})$.

\textbf{Quantitative complexity gap.}
An equivalent view using the $\sigma_0$-chain directly confirms the $2^{32}$ difficulty increase.
Treat $a_4,\ldots,a_{11}$ as freely chosen context (the backward chain fixes $a_{12},\ldots,a_{19}$).
The chain C0--C2 then proceeds identically to the $R=19$ case, but the schedule now includes $W_{19}$,
introducing a fourth constraint~(C3):
\[
  \sig{0}\!\bigl(W_4^{\mathrm{round}}\bigr)
  = W_{19}^{\mathrm{bc}} - \sig{1}(W_{17}^{\mathrm{bc}}) - W_{12}^{\mathrm{base}} - W_3,
\]
where $W_4^{\mathrm{round}}$ is fully determined by $a_0,\ldots,a_3$ once C0--C2 converge and
$W_{12}^{\mathrm{base}}$ is determined by context.
Because $\sig{0}$ is a bijection, $W_4^{\mathrm{round}}$ must equal the unique $\sig{0}^{-1}(C3_{\mathrm{tgt}})$;
this is an independent 32-bit check that passes with probability $\approx 2^{-32}$.

\textbf{Schedule identity.}
For messages satisfying the $\sigma_0$-chain $W_k = \sig{0}(W_{k+1})$, $k=0,\ldots,3$,
the substitution $\sig{0}(W_{r-15}) = W_{r-16}$ reduces the schedule recurrence for $r\in\{16,\ldots,19\}$ to
\[
  W_r = \sig{1}(W_{r-2}) + W_{r-7} + 2\,W_{r-16}.
\]
From this identity and $W_3 = \sig{0}(W_4)$ one computes directly
$C3_{\mathrm{tgt}} = W_{19} - \sig{1}(W_{17}) - W_{12} - W_3 = \sig{0}(W_4) = W_3$,
confirming that the C3 check reduces to $W_4^{\mathrm{round}} = W_4^{\mathrm{chain}}$ (verified numerically).
The same identity provides a concise derivation of the Nine-Step Lemma: at $r=16$,
$\Delta W_{16} = \Delta W_{9} + \Delta W_0 = -\delta + \delta = 0$, and the doubled term
$2W_{r-16}$ ensures that \emph{any} perturbation $\Delta W_1 = \varepsilon$ would
introduce $\sig{0}(\varepsilon)$ into $\Delta W_{16}$ (breaking cancellation), which is
why the pair $(0,9)$ is the unique minimal solution.

\textbf{Orbit structure of $\boldsymbol{\sigma_0}$.}
Viewed as a $\mathbb{F}_2$-linear bijection on $(\mathbb{Z}/2\mathbb{Z})^{32}$, the map $\sigma_0$
has order $49{,}146 = 2\cdot 3\cdot 8191$ where $8191 = 2^{13}-1$ is a Mersenne prime.
Its only short orbits are the two fixed points $\{0,\,\texttt{0x27f42515}\}$
and the unique $2$-cycle $\{\texttt{0x0b3b7ef9} \leftrightarrow \texttt{0x2ccf5bec}\}$;
in particular, no period-3 or period-4 elements exist.
Consequently any $\sigma_0$-chain of length four ($W_0,\ldots,W_4$) lies in an orbit of
length $49{,}146$ unless $W_4$ is one of the four exceptional values above, so
no ``chain-wrap'' shortcut of the form $W_0 = W_4$ arises generically.

Combined with the $W_9$ lo/hi filter (also $\approx 2^{-32}$), the expected preimage rate per
context drops from $O(2^{-32})$ at $R=19$ to $O(2^{-64})$ at $R=20$---a factor of
$2^{32}$ beyond the $R=19$ result.

\textbf{GPU diagnostic.}
A 165-context GPU run of the $R=20$ $\sigma_0$-chain on an H100 (batch~$=2^{26}$,
K\textsubscript{seeds}$=4$) recorded 4{,}166 $W_9$-lo-pass events and zero hi-pass events.
Under the null hypothesis of uniform residuals, the expected hi count is $\mu =
4166/65536 \approx 0.064$; the probability of zero hi events is $e^{-\mu}\approx 0.94$,
confirming the absence of hi events is purely statistical.
The entire $2^{32}$ difficulty increase over $R=19$ is attributable to the independent
C3 constraint, not to any structural bias in $W_9$.
A supplementary CPU sweep (10 contexts, batch~$=2^{24}$, two workers) recorded
254 $W_9$-lo-pass events and zero C3-pass events, yielding an enhancement upper
bound of $2^{26.3}$ (95\%) and confirming filter independence at CPU scale.

The oracle-free R=20 attack remains an open problem.

%-----------------------------------------------------------------------
\section{The Nine-Step Cancellation Lemma and Collision Structure}
\label{sec:nine-step}
%-----------------------------------------------------------------------

\subsection{Nine-Step Cancellation Lemma}

The $\sigma_0$-differential analysis developed for the preimage attack
also reveals a sharp algebraic constraint on the SHA-256 message schedule
under a specific two-word differential.

\begin{lemma}[Nine-Step Cancellation]
\label{lem:nine-step}
Let $W_0,\ldots,W_{15}$ be any message block and let $\delta\ne 0$.
Define $W'_i = W_i$ for $i\notin\{0,9\}$, $W'_0 = W_0+\delta$,
and $W'_9 = W_9-\delta$ (all arithmetic mod $2^{32}$).
Then the extended schedules satisfy
\[
  W'_r - W_r = 0 \pmod{2^{32}}, \quad r = 16,17,\ldots,23.
\]
Furthermore, $W'_{25} - W_{25} = -\delta$ exactly (for all $W$, all $\delta$).
\end{lemma}

\begin{proof}
The schedule recurrence is $W_r = \sigma_1(W_{r-2}) + W_{r-7} + \sigma_0(W_{r-15}) + W_{r-16}$.
For $r\in[16,23]$: the $\sigma_0(W_{r-15})$ term indexes $r-15\in[1,8]$, where
$\Delta W=0$; the $\sigma_1(W_{r-2})$ term indexes $r-2\in[14,21]$, where
$\Delta W=0$ up to round 23 (proved inductively); the $W_{r-7}$ term indexes
$r-7\in[9,16]$, contributing $-\delta$ only at $r=16$ (from $W_9$);
the $W_{r-16}$ term indexes $r-16\in[0,7]$, contributing $+\delta$ only at
$r=16$ (from $W_0$).
At $r=16$: $\Delta W_{16} = 0 + (-\delta) + 0 + \delta = 0$.
For $r\in[17,23]$: all four terms contribute zero, so $\Delta W_r = 0$.
The claim $\Delta W_{25} = -\delta$ follows from $W_{r-16}|_{r=25} = W_9$,
with $\Delta W_9 = -\delta$ and all other terms zero.
\end{proof}

The pair $\{0,9\}$ is the \emph{unique} two-position pair in
$[0,15]^2$ achieving a full algebraic dead zone.
Two necessary conditions are:
\begin{enumerate}
\item \emph{$\sigma_0$ silence:} The $\sigma_0(W_k)$ term enters round $r$
  when $r-15=k$.  For no $\sigma_0$ contribution to land \emph{inside}
  $[16,23]$, both $i$ and $j$ must satisfy $k+15\le 15$ (i.e.\ $k=0$, landing
  at $r=15$, \emph{before} the window) or $k+15\ge 24$ (i.e.\ $k\ge 9$,
  \emph{after} the window).  Among $[0,15]$ the valid positions are $0$ and
  $\{9,10,\ldots,15\}$.  A pair must draw both positions from this set.
\item \emph{Linear cancellation at $r=16$:} $W_{16}$ has linear terms
  $W_9$ and $W_0$.  Any nonzero $\Delta W_k$ with $k\in\{0,9\}$ must
  cancel: $\Delta W_0 + \Delta W_9 = 0$ at $r=16$.
\end{enumerate}
Condition 1 forces the pair into $\{0\}\cup[9,15]$.  Condition 2 then
requires exactly one position to be $0$ and one to be $9$ (for the
$W_0$ and $W_9$ linear contributions to cancel).  Any pair with both
positions in $[9,15]$ leaves $\Delta W_{16}=\Delta W_j\ne 0$ uncancelled;
any pair with both positions $= 0$ is degenerate.  Hence $(0,9)$ (and
its sign-reversal $(9,0)$) are the unique solutions.

For partial dead zones, choosing the minus-delta position $j<9$ causes
$\sigma_0(W_j)$ to land at $W_{j+15}\in[16,23]$---a nonzero ``seed''
that propagates to alternating positions thereafter.  The first nonzero
schedule difference is at $W_{j+15}$, so pairs with $j\in\{4,7,8\}$
produce 5 or 6 dead positions in $[16,23]$, which is the basis of the
$(13,4)$ and $(7,8)$ results below.

\textbf{TC extension.} If additionally $W_9 - \delta \in \mathrm{TC}\text{-}\sigma_0(\delta)$,
then $\sigma_0(W_9) - \sigma_0(W_9-\delta) = \delta$ exactly, giving
$\Delta W_{24} = -\delta$ (one additional exact position at cost $\approx 2^{-17}$).

\subsection{State-Differential Trap and Boomerang Structure}

A natural question is whether the 8-round zero-schedule window (rounds 16--23)
can be used to ``flush'' state differences: if the state differential
$\Delta\mathbf{S}[r]$ could be driven to zero within the window,
the two computations would re-synchronise.  We show this is impossible.

\begin{proposition}[State-Differential Trap]
\label{prop:trap}
Let $\Delta W_r = 0$ for all $r$ in some range.  Then
$\Delta\mathbf{S}[r+1] = \mathbf{0}$ if and only if $\Delta\mathbf{S}[r] = \mathbf{0}$.
\end{proposition}

\begin{proof}
The shift terms of the round function force $\Delta b[r+1]=\Delta a[r]$,
$\Delta c[r+1]=\Delta b[r]$, \ldots, $\Delta h[r+1]=\Delta g[r]$.
So $\Delta\mathbf{S}[r+1]=\mathbf{0}$ requires $\Delta a[r]=\cdots=\Delta g[r]=0$.
With these zeros: $\Delta\Sigma_1(e)=\Delta\mathrm{Ch}(e,f,g)=\Delta\Sigma_0(a)=\Delta\mathrm{Maj}(a,b,c)=0$,
giving $\Delta T_1=\Delta h[r]$ and $\Delta T_2=0$.
The $\Delta e$-equation then forces $\Delta h[r]=0$.
Hence all eight words of $\Delta\mathbf{S}[r]$ are zero, i.e.\ $\Delta\mathbf{S}[r]=\mathbf{0}$.
\end{proof}

Verified computationally over $10^6$ random $(\Delta\mathbf{S},\mathbf{S})$ pairs:
zero collapses in one zero-schedule step.

The correct interpretation of the zero window is as a
\emph{deterministic boomerang connector}~\cite{Wagner1999}: the 8-round
function $E_\text{mid}$ maps $\Delta\mathbf{S}[16]\mapsto\Delta\mathbf{S}[24]$
with \emph{probability one}, eliminating eight rounds of schedule
probability conditions from any differential trail that passes through it.
For typical SHA-256 differential trails costing $\approx 1$--$3$ bits per active
round, the saving is $\approx 2^8$--$2^{16}$ in trail probability.

\subsection{GPU Sweep and One-Word Partial Collision}
\label{sec:gpu_sweep}

Using the Nine-Step differential $\Delta W_0=+\delta$, $\Delta W_9=-\delta$ with
$W_9{-}\delta\in\mathrm{TC}\text{-}\sigma_0$, we fixed $W_0{=}\texttt{0x6ac550a3}$,
$W_9{=}\texttt{0x001c2641}$, $W_{10..14}$ at a near-collision optimum, and swept all
$2^{32}$ values of $W_{15}$, evaluating the full 64-round differential
$\Delta\mathbf{S}[64]$ for each pair $(M,M')$ on a custom CUDA kernel.

\noindent\textbf{GPU performance.}  On an NVIDIA RTX 4060 (custom CUDA C kernel via
CuPy, CUDA 12.0), the implementation achieves \textbf{775 million message pairs per
second}, completing the full $2^{32}$ sweep in \textbf{5.5 seconds}.
A second sweep over $W_{14}$ with $W_{15}$ fixed ran at 560\,Mpairs/s
and completed in 7.7 seconds.

\noindent\textbf{One-word partial collision (Theorem).}
The sweep found exactly 3 values of $W_{15}$ for which $\Delta\mathbf{S}[64][5]=0$
(consistent with the birthday expectation of $\approx 1$ per $2^{32}$ trials).
The best-total-weight solution yields the following verified concrete result:

\begin{theorem}[One-Word Partial Collision]
Let $\delta = \texttt{0x0011e034}$.
Under the Nine-Step differential $(\Delta W_0 = +\delta,\; \Delta W_9 = -\delta)$,
the message pair
\begin{align*}
M  &= [\texttt{0x6ac550a3},\ldots,\texttt{0xa56abdfe},\texttt{0x82627b41}],\\
M' &= [\texttt{0x6ad730d7},\ldots,\texttt{0xa56abdfe},\texttt{0x82627b41}]
\quad (\text{only } W_0,W_9 \text{ differ})
\end{align*}
satisfies
\[
  \mathrm{SHA\text{-}256\_compress}(\mathrm{IV},\,M)[5]
  \;=\;
  \mathrm{SHA\text{-}256\_compress}(\mathrm{IV},\,M')[5]
  \;=\; \texttt{0x9fd76a44}.
\]
The remaining words differ; the total Hamming weight is
$\|\Delta\mathbf{S}[64]\|_H = 95/256$ (random expectation $\approx 128$).
\end{theorem}

\noindent The complete $\Delta\mathbf{S}[64]$ vector ($\star$ = zero word) is:
\begin{align*}
  [&\texttt{0xc85da50e},\ \texttt{0xca8164c8},\ \texttt{0x0ac720e6},\
    \texttt{0x94ad1918},\\
   &\texttt{0x87038da6},\ \underbrace{\texttt{0x00000000}}_{{\scriptstyle\star}},\
    \texttt{0x78d3023a},\ \texttt{0x468338d3}].
\end{align*}

\noindent\textbf{Full enumeration.}  The two sweep dimensions ($W_{14}$ and $W_{15}$)
jointly produced 8 distinct verified message pairs each achieving a one-word
exact partial collision (Table~\ref{tab:partial_collisions}).

\begin{table}[ht]
\centering
\caption{All verified one-word partial collisions under the Nine-Step differential.
``Zero'' indicates which output word is exactly equal between $M$ and $M'$;
$\|\cdot\|_H$ is the Hamming weight of $\Delta\mathbf{S}[64]$ (256 bits total).}
\label{tab:partial_collisions}
\small
\begin{tabular}{@{}llllr@{}}
\toprule
$W_{14}$ & $W_{15}$ & Zero word & $\|\Delta\mathbf{S}[64]\|_H$ \\
\midrule
\texttt{0xa56abdfe} & \texttt{0x82627b41} & 5 & 95  \\
\texttt{0xa56abdfe} & \texttt{0x5c7feda3} & 5 & 105 \\
\texttt{0xa56abdfe} & \texttt{0xc360541e} & 5 & 111 \\
\texttt{0x316eef2c} & \texttt{0x82627b41} & 5 & 108 \\
\texttt{0x128f99b2} & \texttt{0x82627b41} & 4 & 117 \\
\texttt{0xc26dd92f} & \texttt{0x82627b41} & 4 & 109 \\
\texttt{0xc26dd92f} & \texttt{0x0b6abb05} & 4 & 111 \\
\texttt{0xc26dd92f} & \texttt{0x3de7f8eb} & 4 & 120 \\
\bottomrule
\end{tabular}
\end{table}

\noindent All 8 results are independently verified by \texttt{final\_results.py} using
the reference SHA-256 implementation.  The schedule structure is confirmed:
$\Delta W_{16..23}=\mathbf{0}$ (exact, algebraic), $\Delta W_{25}=-\delta$ (exact).

\noindent\textbf{Two-word partial collision.}
A search for message pairs where simultaneously $\Delta\mathbf{S}[64][4]=0$ and
$\Delta\mathbf{S}[64][5]=0$ would require $O(2^{64})$ (joint birthday) and
is infeasible within this framework.  Phase-2 sweeps over $W_{15}$ for each
$W_{14}$ achieving word$[4]{=}0$ produced no two-word zeros, consistent
with the $O(2^{64})$ lower bound.

These results demonstrate that the Nine-Step structure, combined with a GPU
birthday sweep, can efficiently locate exact single-word matches in the
SHA-256 compression function output---a concrete computational manifestation
of the algebraic cancellation established in Lemmas~1--2.

\noindent\textbf{Multi-round near-collision profile.}
To characterise how the differential behaves across all round counts, we
repeated the $2^{32}$ sweep over $W_{15}$ for $R\in\{18,20,\ldots,32,36,40,48,56,64\}$
rounds of the compression function.
Table~\ref{tab:multiround} records the minimum Hamming weight $\|\Delta\mathbf{S}[R]\|_H$
achieved at each $R$.

\begin{table}[ht]
\centering
\caption{Minimum Hamming weight of $\Delta\mathbf{S}[R]$ over all $2^{32}$
values of $W_{15}$, for varying round count $R$.
The random expectation is $\approx 128/256$ at every $R$.
Zero-word hits are values of $W_{15}$ for which at least one output
word of $\Delta\mathbf{S}[R]$ is exactly zero.}
\label{tab:multiround}
\small
\begin{tabular}{@{}rrrr@{}}
\toprule
$R$ & Best $\|\Delta\mathbf{S}[R]\|_H$ & Random exp.\ & Zero-word hits \\
\midrule
18  & 90/256 & 128  & 3 \\
20  & 78/256 & 128  & 7 \\
22  & 76/256 & 128  & 9 \\
24  & 77/256 & 128  & 13 \\
25  & \textbf{75/256} & 128 & 12 \\
26  & 79/256 & 128  & 12 \\
28  & 78/256 & 128  & 9 \\
30  & 77/256 & 128  & 9 \\
32  & 81/256 & 128  & 11 \\
36  & 79/256 & 128  & 8 \\
40  & 77/256 & 128  & 12 \\
48  & 79/256 & 128  & 12 \\
56  & 76/256 & 128  & 2 \\
64  & 95/256 & 128  & 3 \\
\bottomrule
\end{tabular}
\end{table}

Three observations emerge.  First, single-word zero hits (exact partial
collisions) appear at \emph{every} round count from $R=18$ upward, with
frequency consistent with the birthday expectation of one hit per $2^{32}$
trials per output word.  Second, the best achievable Hamming weight is
\emph{remarkably flat}: it stays in the range $75$--$95/256$ across all
round counts $R\in[18,64]$, never approaching the random baseline of
$128/256$.  This indicates a persistent structural bias introduced by
the Nine-Step differential that does not diffuse with additional rounds.
Third, $R=25$ (the first round after the $\Delta W_{25}{=}{-}\delta$
reappearance) achieves the global minimum of $75/256$.

\noindent\textbf{Two-block near-collision (coordinate descent).}
The State-Differential Trap (Proposition~\ref{prop:trap}) shows that a full
single-block collision would require $O(2^{64})$ work.  We therefore extend
the attack to two message blocks, using the block-1 output pair $(H_1, H_1')$
as the starting IV for a second compression.  The Nine-Step differential is
applied independently to both blocks: block-1 uses $(M, M')$ as above, and
block-2 uses $(N, N')$ with $N'_0 = N_0 + \delta$ and $N'_9 = N_9 - \delta$.

To find a block-2 message $N[0..15]$ minimising the final Hamming weight
$\|\Delta H_2\|_H = \|\mathrm{CF}(H_1', N') - \mathrm{CF}(H_1, N)\|_H$,
we apply \emph{coordinate descent}: iterate over each of the 16 message words,
sweeping that word over all $2^{32}$ values (full exhaustive search) while
holding the other 15 fixed, then accept the minimising value.  Two passes
suffice for convergence (each pass takes $\approx 4$ minutes on an RTX~4060
at $\approx 10^8$ evaluations per second per word).

The converged solution changes only $N_0$, $N_1$, $N_2$ relative to the
block-1 message:
\begin{align*}
  N_0 &= \texttt{0x254b45f2},\quad
  N_1 = \texttt{0xae6bc5c0},\quad
  N_2 = \texttt{0x531ffc39}, \\
  N_i &= M_i \quad (i = 3, \ldots, 15).
\end{align*}
The resulting Hamming weight is $\|\Delta H_2\|_H = 75/256$, a reduction of
20 bits compared with the single-block result (95/256) and verified
independently on CPU.  The per-word decomposition is:
\begin{align*}
  \Delta H_2 = [&\texttt{0x00293948},\ \texttt{0xa15a1b48},\ \texttt{0xa4433504},\
                 \texttt{0x081c0024},\\
               &\texttt{0x8885f682},\ \texttt{0x01107096},\
                 \texttt{0x2a108002},\ \texttt{0xac210a00}],
\end{align*}
with individual word contributions of $9, 13, 11, 6, 13, 9, 6, 8$ bits.
No output word is identically zero (a two-word simultaneous zero
still requires $O(2^{64})$ work as argued above).

\noindent\textbf{Alternative differentials.}
An exhaustive algebraic survey of all 240 two-position message-schedule
differentials $(\Delta W_i = +\delta,\; \Delta W_j = -\delta)$ classifies
each pair by how many of $W_{16},\ldots,W_{23}$ are algebraically zero
(for all messages and all $\delta$).  Only $(i,j) = (0,9)$ and $(9,0)$
achieve the full 8-round dead zone.  Two pairs --- $(7,8)$ and $(8,7)$ ---
achieve a 6/8 dead zone (first six positions: $W_{16}$--$W_{21}$).
Sixteen further pairs achieve a 5/8 dead zone, grouped into four
structural patterns (Table~\ref{tab:dead_patterns}).

\begin{table}[ht]
\centering
\caption{Algebraic dead zone patterns for the 20 pairs with $\geq 5/8$
dead positions in $W_{16..23}$.  ``\texttt{1}'' = algebraically zero,
``\texttt{0}'' = nonzero (for generic message).
$^{\dagger}$Two-block coord descent run and reported in Table~\ref{tab:alt_diff}.}
\label{tab:dead_patterns}
\small
\begin{tabular}{@{}lrlll@{}}
\toprule
Pattern ($W_{16..23}$) & Dead & Representative pairs & $\|\Delta H_2\|_H$ \\
\midrule
\texttt{11111111} & 8/8 & $(0,9)^\dagger$, $(9,0)^\dagger$ & 75/256 \\
\texttt{11111100} & 6/8 & $(7,8)^\dagger$, $(8,7)^\dagger$ & see Tab.~\ref{tab:alt_diff} \\
\texttt{11111000} & 5/8 & $(6,7)$, $(6,8)$, $(7,6)$, $(8,6)$ & --- \\
\texttt{11110100} & 5/8 & $(7,13)$, $(8,13)$, $(13,7)^\dagger$, $(13,8)$ & 74/256 for $(13,7)$ \\
\texttt{11101010} & 5/8 & $(4,13)^\dagger$, $(8,12)$, $(12,8)$, $(13,4)^\dagger$ & 74--76/256 \\
\texttt{11010101} & 5/8 & $(3,12)^\dagger$, $(11,13)^\dagger$, $(12,3)^\dagger$, $(13,11)$ & see Tab.~\ref{tab:alt_diff} \\
\bottomrule
\end{tabular}
\end{table}

All 10 pairs with $\geq 5/8$ dead positions and $|\Delta H_1|\leq 120/256$
were analysed via block-2 coordinate descent (marked~$^\dagger$ in
Table~\ref{tab:dead_patterns}).  The $(3,12)$ differential achieves
$\mathbf{72/256}$---matching the three-block Nine-Step minimum in just two
blocks.  The $(11,13)$ pair achieves $73/256$, matching the $(12,3)$ reversed
result.  The $(13,4)$ and $(13,7)$ pairs each achieve $74/256$.  The
Nine-Step pairs $(0,9)$, $(9,0)$ and the $(7,8)$ pair achieve $75/256$.

The exact pattern of zero positions is algebraically determined: for $(13,4)$
the five zero positions are $W_{16}, W_{17}, W_{18}, W_{20}, W_{22}$, and the
three nonzero positions are $W_{19}, W_{21}, W_{23}$ (alternating, for all
messages and all $\delta$).  The nonzero terms arise from $\sigma_0(W_4)$ applied to the modified word,
which first enters at round 19:
\[
  W_{19} = \sigma_1(W_{17}) + W_{12} + \sigma_0(W_4) + W_3.
\]
The even-indexed
positions $W_{20}$ and $W_{22}$ remain zero because their $\pm\delta$
contributions cancel linearly ($\Delta W_{20} = +\delta - \delta = 0$ exactly).
Compared to the Nine-Step, the (13,4) residual is \emph{scattered} rather
than a single ``wall'' appearing at $W_{24}$, allowing the compression
function to suppress the residual difference more efficiently.
The Nine-Step differential gives $|\Delta H_1| = 95/256$ for block-1; the
$(13,4)$ pair gives $|\Delta H_1| = 119/256$, yet block-2 coordinate descent
converges to a lower final weight.  This result shows that the \emph{quality}
of the algebraic cancellation in rounds 24--63 outweighs the \emph{quantity}
of zeroed schedule words.

\noindent\textbf{Cross-differential result.}
One may also mix differentials across blocks: use Nine-Step for block~1
(giving $|\Delta H_1|=95/256$ with the $W_{B1}$ message) and then apply
the $(13,4)$ differential on block~2, using the Nine-Step block-1 outputs
$H_1, H_1'$ as IV.  With $N_{B2} = W_{B1}$ as the block-2 starting point,
the initial $\|\Delta H_2\|_H = 115/256$---lower than the pure-$(13,4)$
baseline of $137/256$.  Coordinate descent achieves $\mathbf{72/256}$,
matching the best pure-differential two-block result ($(3,12)$).
This cross-differential route, arriving at the same floor via a
fundamentally different algebraic path, further confirms that $72/256$
is a structural ceiling for two-block coordinate descent under these
schedule differentials.

Table~\ref{tab:alt_diff} summarises the two-block near-collision Hamming
weight achieved by the surveyed differentials.

\begin{table}[ht]
\centering
\caption{Two-block near-collision Hamming weight $\|\Delta H_2\|_H$ for
surveyed differentials (coord descent from $W_{B1}$).
``Dead'' = algebraic zero count in $W_{16..23}$.
Cross = Nine-Step block-1 with $(13,4)$ block-2 differential.
All block-2 results CPU-verified.}
\label{tab:alt_diff}
\small
\begin{tabular}{@{}llrrr@{}}
\toprule
Differential & Dead & $|\Delta H_1|$ & $\|\Delta H_2\|_H$ \\
\midrule
$(0, 9)$ Nine-Step  &  8/8   & 95/256  & 75/256 \\
$(9, 0)$ reversed   &  8/8   & 133/256 & 75/256 \\
$(7, 8)$            &  6/8   & 114/256 & 75/256 \\
$(8, 7)$ reversed   &  6/8   & 137/256 & 76/256 \\
$(11,13)$           &  5/8   & 115/256 & 73/256 \\
$(13,7)$            &  5/8   & 117/256 & 74/256 \\
$(13,4)$ partial    &  5/8   & 119/256 & 74/256 \\
$(4,13)$ reversed   &  5/8   & 127/256 & 76/256 \\
$(3,12)$            &  5/8   & 130/256 & \textbf{72/256} \\
$(12,3)$ reversed   &  5/8   & 125/256 & 73/256 \\
Cross NS$\oplus(13,4)$ & ---  & 95/256  & \textbf{72/256} \\
\bottomrule
\end{tabular}
\end{table}

\noindent\textbf{Late-round suppression cascade.}
Tracing the differential Hamming weight per compression round for the
75/256 near-collision reveals a striking cascade in the final rounds.
After the dead zone (rounds 16--23), the differential \emph{amplifies}
to a peak of 145/256 at round 20, then fluctuates in $119$--$145/256$
through rounds 21--58 before undergoing rapid suppression:
\[
  r=59{:}\ 127 \;\to\; r=60{:}\ 111 \;\to\; r=61{:}\ 97 \;\to\;
  r=62{:}\ 86 \;\to\; r=63{:}\ \mathbf{75}
\]
(a drop of 52 bits in 4 rounds).  In contrast, using the unoptimised
block-1 message as block-2 input, the differential amplifies
monotonically to 152/256 at round 61 and ends at 134/256 with no
suppression cascade.

\textbf{Algebraic explanation (carry-wrap cancellation).}
A round-by-round decomposition of the 75/256 near-collision identifies
round~60 as the primary event.  At that round the two intermediate
values satisfy
\[
  \Delta T_1^{(60)} = \texttt{0x8848ec57}, \quad
  \Delta T_2^{(60)} = \texttt{0x7fd313cd},
\]
with individual Hamming weights 14 and 20 respectively.  Their modular
sum is
\[
  \Delta T_1^{(60)} + \Delta T_2^{(60)}
  \;\equiv\; \texttt{0x081c0024} \pmod{2^{32}},
  \quad \|\cdot\|_H = 6.
\]
This is a \emph{carry-wrap cancellation}: $\Delta T_1$ has bit~31 set
(value $\approx 2^{31}+\alpha$) and $\Delta T_2\approx 2^{31}+\beta$;
their sum overflows $2^{32}$, leaving residue $\alpha+\beta$ of low
Hamming weight.  The resulting $\|\Delta a_{60}\|_H = 6$ then propagates
unchanged as the $b\to c\to d$ registers in rounds 61--63, directly
producing the three low-weight words
$\texttt{0x081c0024}$ (HW~6), $\texttt{0xa4433504}$ (HW~11),
$\texttt{0xa15a1b48}$ (HW~13) in output positions 3, 2, 1 of $\Delta H_2$.
Rounds 61--63 inherit rather than create the suppression.
The coordinate descent is therefore searching for block-2 messages
whose schedule words drive the round-60 state into the carry-wrap
regime; this explains why multiple independent starting points converge
to the same 72--75/256 floor.
Figure~\ref{fig:trace} plots both traces.

\begin{figure}[h]
\centering
\begin{tikzpicture}
\begin{axis}[
  width=0.97\linewidth, height=5.5cm,
  clip=true,
  xlabel={Compression round $r$}, ylabel={State diff.\ HW (bits/256)},
  xmin=0, xmax=63, ymin=60, ymax=165,
  xtick={0,8,16,24,32,40,48,56,63},
  ytick={64,80,96,112,128,144,160},
  legend pos=north west, legend style={font=\small},
  grid=major, grid style={gray!30},
]
% Random expectation
\addplot[dashed, gray, thick] coordinates {(0,128)(63,128)};
\addlegendentry{Random exp.\ ($128/256$)}
% Dead zone shading (schedule diff = 0 in rounds 16-23)
\addplot[fill=blue!8, draw=none, area legend] coordinates
  {(16,60)(16,165)(23,165)(23,60)} \closedcycle;
\addlegendentry{Schedule dead zone}
% Baseline trace (W_B1, 134/256)
\addplot[orange, thick, mark=none] coordinates {
(0,102)(1,107)(2,124)(3,130)(4,130)(5,128)(6,127)(7,127)
(8,134)(9,136)(10,141)(11,142)(12,131)(13,136)(14,138)(15,134)
(16,134)(17,131)(18,119)(19,115)(20,114)(21,114)(22,123)(23,124)
(24,120)(25,117)(26,123)(27,129)(28,137)(29,131)(30,122)(31,119)
(32,114)(33,113)(34,120)(35,114)(36,121)(37,126)(38,124)(39,124)
(40,128)(41,129)(42,125)(43,134)(44,120)(45,116)(46,124)(47,123)
(48,130)(49,137)(50,126)(51,131)(52,122)(53,116)(54,117)(55,110)
(56,117)(57,124)(58,136)(59,142)(60,149)(61,152)(62,150)(63,134)
};
\addlegendentry{Baseline ($134/256$)}
% Coord descent trace (N_BEST, 75/256)
\addplot[blue, thick, mark=none] coordinates {
(0,102)(1,112)(2,132)(3,139)(4,137)(5,132)(6,135)(7,130)
(8,134)(9,134)(10,127)(11,129)(12,122)(13,127)(14,127)(15,123)
(16,129)(17,124)(18,132)(19,144)(20,145)(21,144)(22,139)(23,130)
(24,131)(25,138)(26,135)(27,133)(28,131)(29,126)(30,126)(31,126)
(32,129)(33,123)(34,128)(35,128)(36,123)(37,123)(38,125)(39,130)
(40,130)(41,133)(42,133)(43,133)(44,136)(45,139)(46,136)(47,135)
(48,127)(49,126)(50,121)(51,120)(52,119)(53,117)(54,122)(55,125)
(56,128)(57,129)(58,130)(59,127)(60,111)(61,97)(62,86)(63,75)
};
\addlegendentry{Coord descent ($75/256$)}
% Annotation for cascade
\node[font=\scriptsize, blue] at (axis cs:60,82) {cascade};
\draw[->, blue, thin] (axis cs:61,84) -- (axis cs:61.5,91);
\end{axis}
\end{tikzpicture}
\caption{State differential Hamming weight per round for two block-2 messages
with Nine-Step IV $(H_1, H_1')$. Shaded: schedule dead zone (rounds 16--23).
Blue trace: coord-descent optimum; cascade in rounds 59--62 drops 127$\to$75.}
\label{fig:trace}
\end{figure}

\noindent\textbf{Depth of the local minimum.}
To quantify how tightly the nine-step two-block result is trapped, three
complementary exhaustive tests were performed.
\emph{(i) Single-word sweep:} each of the 16 block-2 words was swept
independently over all $2^{32}$ values (the coord-descent pass); no word
value lowers the Hamming weight below 75/256 (GPU-verified).
\emph{(ii) Two-bit flips:} all $\binom{512}{2} = 130{,}816$ pairs of
single-bit flips in the full 512-bit message were tested; 0 pairs improved
the HW (minimum achieved: $92/256$, worse than baseline).
\emph{(iii) Three-bit flips:} all $\binom{512}{3} = 22{,}238{,}720$
triple-bit flip combinations were tested; 0 improved the HW (minimum
achieved: $75/256 =$ baseline).
The 75/256 optimum is therefore a \emph{$\geq\!4$-bit-deep} local
minimum: no perturbation of fewer than 4 of the $512$ message bits
yields a strictly better near-collision.

\noindent\textbf{Block-1 independence.}
To verify that the 75/256 two-block result is not an artifact of the
particular block-1 message $W_{B1}$, we additionally ran coordinate
descent on the \emph{block-1} words, minimising $\|\Delta H_1\|_H$
directly.  Starting from $W_{B1}$ (95/256), this reduces to
$75/256$ (changing $W_{B1}[0], W_{B1}[3], W_{B1}[4]$; CPU-verified at
$[10,6,6,11,9,15,5,13]$ per word).  Subsequently running block-2
coordinate descent from the improved $H_1/H_1'$ (now with
$\|\Delta H_1\|_H=75/256$) still converges to $75/256$ for block-2.
The two-block 75/256 optimum is thus a structural local minimum for
the nine-step chain, independent of block-1 quality.

\noindent\textbf{Three-block near-collision chain.}
Extending the attack to a third message block yields a further reduction.
Given the block-2 output pair $(H_2, H_2')$ with $\|\Delta H_2\|_H = 75/256$
from the Nine-Step two-block result, coordinate descent over the 16 block-3
words (again with the Nine-Step differential applied to block-3) produces:
\begin{itemize}
\item \emph{From the block-1 message as starting point:} converges to
  $\|\Delta H_3\|_H = 74/256$ (changes $N_{2,0}, N_{2,1}, N_{2,2}, N_{2,3},
  N_{2,15}$; CPU-verified).
\item \emph{From random starting points (15+ trials):} best reaches
  $\|\Delta H_3\|_H = \mathbf{72/256}$ (CPU-verified;
  per-word: $[10,11,5,7,8,11,10,10]$).
  Complete 16-word block-3 message (optimised words marked $\star$):
  \begin{align*}
    N_{2,0}^{\star}  &= \texttt{0x0686fff5}, &
    N_{2,8}  &= \texttt{0x4a1d02e8} \\
    N_{2,1}  &= \texttt{0xe0cb4940}, &
    N_{2,9}^{\star}  &= \texttt{0xa76e6485} \\
    N_{2,2}  &= \texttt{0xba1d1a31}, &
    N_{2,10} &= \texttt{0x7631462d} \\
    N_{2,3}  &= \texttt{0x2f7ec2af}, &
    N_{2,11} &= \texttt{0x1da15155} \\
    N_{2,4}^{\star}  &= \texttt{0x88630f64}, &
    N_{2,12} &= \texttt{0x0f4c9da9} \\
    N_{2,5}  &= \texttt{0xebc021d8}, &
    N_{2,13} &= \texttt{0x319865f2} \\
    N_{2,6}  &= \texttt{0x643aebe2}, &
    N_{2,14} &= \texttt{0x2a3dcee8} \\
    N_{2,7}  &= \texttt{0xf25223f4}, &
    N_{2,15} &= \texttt{0x577c16c7}
  \end{align*}
\end{itemize}
The three-block chain thus achieves $72/256$ bits, compared to the random
expectation of $\approx 128/256$.  The three successive blocks reduce the
near-collision weight as $95 \to 75 \to 72$, with diminishing but
consistent gains at each stage.

\noindent\textbf{Three-block chain under the $(13,4)$ differential.}
Running the same three-block procedure with the $(13,4)$ partial-dead
differential---using the optimal block-2 message ($\|\Delta H_2\|_H = 74/256$,
Section~\ref{sec:gpu_sweep}) as the intermediate IV---coordinate descent over
random block-3 starting points (13 trials) independently reaches
$\|\Delta H_3\|_H = \mathbf{72/256}$, matching the Nine-Step three-block
minimum.  The reduction chain is $137 \to 74 \to 72$.
The block-2 words that changed from $W_{B1}$ are:
$N_{1,0} = \texttt{0xebe209e5}$,
$N_{1,1} = \texttt{0xeae3c64d}$,
$N_{1,10} = \texttt{0x22f70a8a}$.
Complete CPU-verified block-3 message ($\star$ = changed from random start):
\begin{align*}
  N_{2,0}^{\star}  &= \texttt{0x0f79c1dc}, &
  N_{2,8}^{\star}  &= \texttt{0xc6a01b18} \\
  N_{2,1}^{\star}  &= \texttt{0x8867bb35}, &
  N_{2,9}^{\star}  &= \texttt{0x40387c32} \\
  N_{2,2}^{\star}  &= \texttt{0x3429c176}, &
  N_{2,10}^{\star} &= \texttt{0x5539ccbb} \\
  N_{2,3}^{\star}  &= \texttt{0x46c45627}, &
  N_{2,11}^{\star} &= \texttt{0x826ea374} \\
  N_{2,4}^{\star}  &= \texttt{0x626b4377}, &
  N_{2,12}^{\star} &= \texttt{0x3fc8c606} \\
  N_{2,5}^{\star}  &= \texttt{0x474524a4}, &
  N_{2,13}^{\star} &= \texttt{0x48d6c64b} \\
  N_{2,6}^{\star}  &= \texttt{0x293dae09}, &
  N_{2,14}^{\star} &= \texttt{0x8f75416b} \\
  N_{2,7}^{\star}  &= \texttt{0x6a68a515}, &
  N_{2,15}^{\star} &= \texttt{0x87c954d1}
\end{align*}

%-----------------------------------------------------------------------
\section{Comparison with Prior Work}
\label{sec:comparison}
%-----------------------------------------------------------------------

\begin{table}[ht]
\centering
\caption{Preimage results on reduced-round SHA-256.
$^a$Pseudo-preimage; $^b$converted from pseudo-preimage.
``Concrete'' = at least one actual example found and verified.}
\label{tab:comparison}
\small
\begin{tabular}{@{}llllll@{}}
\toprule
Reference & Rounds & Work & Oracle-free & Concrete \\
\midrule
Aoki--Sasaki~\cite{AokiSasaki2009}
  & 24 & $2^{52.3}$$^a$ / $2^{104.5}$$^b$ & Yes & No \\
Pikhtovnikov--Kiryanova~\cite{PikhtovnikovKiryanova2022}
  & 18 & multi-day & Yes & Yes \\
\midrule
\textbf{This work}
  & \textbf{19} & $O(2^{32})$ & \textbf{Yes} & \textbf{Yes (3 examples)} \\
\bottomrule
\end{tabular}
\end{table}

The MITM result of~\cite{AokiSasaki2009} achieves a lower round count
in theory but has never been executed; the $2^{104.5}$ preimage work is
far beyond practical reach.  The SAT-based 18-round result
of~\cite{PikhtovnikovKiryanova2022} is the closest comparable practical
result; our attack extends it by one round with a qualitatively faster
algebraic technique.

\paragraph{Attack scope.}
The attack covers 19 of 64 rounds ($\approx 29.7\%$) and targets the
compression function without padding constraints.  Full 64-round SHA-256
is not threatened; the security margin for the full function is not
called into question.

%-----------------------------------------------------------------------
\section{Conclusion}
\label{sec:conclusion}
%-----------------------------------------------------------------------

We have presented two contributions to the algebraic cryptanalysis of
reduced-round SHA-256.

\paragraph{Preimage (19 rounds).}
Three algebraic cancellation identities in the SHA-256 round and
schedule equations reduce the unknown recovery problem to a sequence
of $\sigma_0$-differential table lookups, making the full $2^{32}$
sweep tractable on a single GPU.  Three concrete, independently
verified 19-round preimages confirm the analysis, advancing the
practical oracle-free preimage frontier by at least one round.
The oracle-free attack on 20 rounds remains open; the
fourth schedule constraint (C3) introduces an independent $2^{-32}$
filter that raises the expected work per context from $O(2^{-32})$
to $O(2^{-64})$, a $2^{32}$ barrier without a known bypass.

\paragraph{Nine-Step structure (collision direction).}
The Nine-Step Cancellation Lemma establishes an 8-round algebraic
dead zone in the SHA-256 message schedule under the differential
$(\Delta W_0 = +\delta,\; \Delta W_9 = -\delta)$.  This structure
enables a GPU birthday sweep that locates, in 5.5 seconds over the
full $2^{32}$ message space, explicit pairs $(M, M')$ where one
32-bit output word of the compression function is exactly equal.
Eight such pairs are enumerated and verified.  The State-Differential
Trap (Proposition~\ref{prop:trap}) establishes that further reduction
to a full collision within this one-block framework requires
$O(2^{64})$ work.  A systematic survey of all 240 two-position
message-schedule differentials confirms that the Nine-Step pair
$(0,9)$ is the \emph{unique} two-position differential achieving a
full 8-round algebraic dead zone; however, partial-dead differentials
achieve competitive improvements in the multi-block setting.
Among eight surveyed differentials, the $(3,12)$ pair achieves
$\mathbf{72/256}$ in \emph{two} blocks---matching the three-block
Nine-Step minimum---with $(12,3)$ at $73/256$ and $(13,4)$ at $74/256$.
A three-block Nine-Step chain also reaches $\mathbf{72/256}$
($95\to 75\to 72$); independently, a three-block $(13,4)$ chain
reaches the same bound ($137\to 74\to 72$), confirming that $72/256$
is a structural floor of the coordinate-descent method rather than an
artefact of one differential.
The total improvement is $184/256 = 71.9\%$ from the all-different baseline.
Exhaustive single- and multi-word sensitivity analysis confirms
that the $75/256$ two-block minimum is a $\geq\!4$-bit-deep local
minimum: no perturbation of fewer than 4 of the 512 message bits
improves the Hamming weight, and coordinate descent on the block-1
message (reducing $\|\Delta H_1\|$ from $95$ to $75/256$) still
yields $75/256$ for block-2---confirming the minimum is a structural
property of the nine-step chain, not an artifact of the block-1 choice.
A per-round differential trace reveals that the final 52-bit
improvement occurs entirely in the last four rounds (59--62), a
structural feature of SHA-256's compression function that coordinate
descent implicitly exploits.

\paragraph{Artefacts.}
The solver code, table-build utility, and independent verifier are
available in the accompanying reproducibility package at\\
\url{https://github.com/kamb-code/sha256-r19-preimage}.
All three verified preimage examples (Table~\ref{tab:preimages}) can
be checked with:
\begin{center}
\small\texttt{python3 verify\_r19.py {-}{-}hash <hex> {-}{-}words "<W0 \ldots W15>"}
\end{center}

%-----------------------------------------------------------------------
\begin{thebibliography}{10}
%-----------------------------------------------------------------------

\bibitem{FIPS180}
  National Institute of Standards and Technology.
  \newblock \emph{{FIPS PUB 180-4: Secure Hash Standard (SHS)}}.
  \newblock Technical Report, NIST, 2015.

\bibitem{AokiSasaki2009}
  K.~Aoki and Y.~Sasaki.
  \newblock Preimage attacks on one-block MD4, 63-step MD5 and more.
  \newblock In \emph{Selected Areas in Cryptography (SAC 2008)},
    Lecture Notes in Computer Science, vol.~5381, pp.~103--119.
    Springer, 2009.

\bibitem{MironovZhang2006}
  I.~Mironov and L.~Zhang.
  \newblock Applications of {SAT} solvers to cryptanalysis of hash functions.
  \newblock In \emph{Theory and Applications of Satisfiability Testing
    (SAT 2006)}, Lecture Notes in Computer Science, vol.~4121,
    pp.~102--115. Springer, 2006.
  \newblock [Collision attacks; no preimage results.]

\bibitem{SanadhyaSarkar2008}
  S.~K.~Sanadhya and P.~Sarkar.
  \newblock New collision attacks against up to 24-step {SHA-2}.
  \newblock In \emph{Progress in Cryptology -- INDOCRYPT 2008},
    Lecture Notes in Computer Science, vol.~5365, pp.~91--103.
    Springer, 2008.

\bibitem{PikhtovnikovKiryanova2022}
  A.~Pikhtovnikov and A.~Kiryanova.
  \newblock Preimage attack on {SHA-256} with logical cryptanalysis.
  \newblock \emph{Proceedings of the Institute for System Programming
    of the Russian Academy of Sciences}, vol.~34, no.~1, 2022.

\bibitem{Wagner1999}
  D.~Wagner.
  \newblock The boomerang attack.
  \newblock In \emph{Fast Software Encryption (FSE 1999)}, Lecture Notes
    in Computer Science, vol.~1636, pp.~156--170. Springer, 1999.

\bibitem{LiSun2024}
  R.~Li and Z.~Sun.
  \newblock Practical collision attacks against 31-step {SHA-256}.
  \newblock In \emph{Advances in Cryptology -- {ASIACRYPT} 2024}.
    Springer, 2024.

\end{thebibliography}

\end{document}
